---
title: "Illustrative Examples of LHmom"
author: "Palakorn Seenoi, Yire Shin, Piyapatr Busababodhin, Jeong-Soo Park"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Illustrative Examples of LHmom}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
This vignette demonstrates the core functionalities of the ``LHmom`` package using the built-in ``bangkok`` dataset. This dataset contains the annual maximum series of daily rainfall in Bangkok, Thailand, from 1951 to 2025.

We will walk through the process of calculating sample LH-moments, estimating parameters for the Generalized Extreme Value (GEV) distribution, and evaluating the model fit.

## 1. Installation
You can install the development version of LHmom from GitHub with:
```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("palakorn-seenoi/LHmom")
```

## 2. Data Preparation
Load the package and the dataset:
```{r}
library(LHmom)
data(bangkok)

# View the first few rows
head(bangkok)
```
## 3. LH estimation for GEV distribution
Users can compute the sample LH-moments for a given order $\eta$ and subsequently estimate the parameters for various distributions. The following example demonstrates fitting the GEV distribution using the first order of LH-moments ($\eta = 1$).

### Calculating Sample LH-moments
```{r}
# Calculate sample LH-moments at eta = 1
sample_lh <- lhmoms(bangkok$rainfall, eta = 1)
print(sample_lh$lambdas)
```
### Parameter Estimation
Next, we estimate the GEV parameters using the calculated sample LH-moments:
```{r}
# Estimate GEV parameters (eta = 1)
fit_gev <- lh.pargev(bangkok$rainfall, eta = 1)
print(fit_gev$para)
```
### Verifying Numerical Consistency
To verify the estimation algorithm, we can calculate the theoretical LH-moments from our newly fitted parameters:
```{r}
# Calculate theoretical LH-moments from the fitted parameters
theo_lh <- lhmom.gev(fit_gev$para, eta = 1)
print(theo_lh$lambdas)
```
Notice the numerical consistency: because the GEV is a three-parameter distribution, the estimation algorithm successfully equates the first three sample LH-moments with their theoretical counterparts.

## 4. Wang's Goodness-of-Fit Test
A crucial aspect of extreme value analysis is model validation. The ``wang.test.lhgev()`` function evaluates whether the GEV distribution provides a suitable fit to the data across different LH-moment orders ($\eta$) by calculating a Z-statistic based on the difference between the sample and theoretical LH-kurtosis.
```{r}
# Perform Wang's goodness-of-fit test
gof_test <- wang.test.lhgev(bangkok$rainfall)
print(gof_test)
```
### Interpretation:
The results demonstrate a satisfactory fit across all $\eta$ values, as all $p$-values significantly exceed the 0.05 significance level. Notably, at $\eta = 3$, the absolute Z-statistic reaches its minimum, and the $p$-value approaches unity. This indicates that employing a higher-order LH-moment ($\eta = 3$) captures the upper-tail distributional shape of the annual maximum rainfall exceptionally well compared to ordinary L-moments ($\eta = 0$).
